Evolution of pre-copulatory and post-copulatory inbreeding behaviour and associated

### 2016-12-15 ms_pre_post.Rmd
Edit up to right before the (soon to be former?) 'Verbal Theory' subsection in the introduction

Inbreeding avoidance has been widely hypothesised to evolve through pre-copulatory and post-copulatory mechanisms as a consequence of inbreeding depression in offspring viability and female multiple mating (polyandry). However, no theoretical framework exists from which to derive hypothese concerning joint evolution of pre-copulatory and post-copulatory inbreeding avoidance. Evolution of post-copulatory inbreeding avoidance necessitates polyandry, potentially leading to complex evolutionary dynamics among inbreeding strategies and female mating frequency, each of which might be affected by direct negative costs. Hence, we use individual-based modelling to quantify evolution of interacting pre- and post-copulatory inbreeding avoidance and associated polyandry across multiple cost combinations given severe inbreeding depression. We find that evolution of post-copulatory inbreeding avoidance increases selection for polyandry, that selection for costly inbreeding strategies was weak or negligible given a non-costly alternative strategy. Established pre-copulatory inbreeding avoidance generally precluded evolution of polyandry and therefore inbreeding avoidance, but established post-copulatory inbreeding aviodance did not preclude pre-copulatory inbreeding avoidance from evolving. We thereby provide a general framework for understanding evolution of pre- and/or post-copulatory inbreeding avoidance across systems, and conclude that evolution of inbreeding avoidance phenotypes and associated polyandry are predicted to interact and be strongly affected by direct costs.

@@

Inbreeding often greatly reduces the fitness of inbred offspring [termed 'inbreeding depression'; @Charlesworth1999; @Keller2002; @Charlesworth2009]. Such strong inbreeding depression is widely hypothesised to drive evolution of inbreeding avoidance, which can be enacted through multiple reproductive decisions in species with obligate biparental reproduction [@Parker1979; @Parker2006; @Pusey1996; @Szulkin2012]. From the perspective of a focal breeding female, inbreeding avoidance might be achieved by avoiding mating with related males (i.e., pre-copulatory inbreeding avoidance), or by biasing fertilisation towards unrelated males after mating has occurred, meaning that related males are less likely to sire offspring (i.e., post-copulatory inbreeding avoidance). Importantly, evolution of post-copulatory inbreeding avoidance requires that females are polyandrous, defined as mating with multiple males during a single reproductive bout. Further, such polyandry could itself evolve because it allows females to mate with additional unrelated males following initial mating with a relative, thereby facilitating pre-copulatory inbreeding avoidance even in the absence of any post-copulatory female choice or otherwise biased fertilisation (e.g., given a 'fair raffle' among sperm provided by females' mates). As a reproductive strategy, polyandry can thereby simultaneously allow females to mate with less closely related males and create the opportunity for further inbreeding avoidance enacted through active post-copulatory choice. Polyandry has consequently been widely hypothesised to evolve as an adaptation to allow females to avoid inbreeding and thereby increase offspring fitness [@Zeh1997; @Jennions2000; @Tregenza2002; @Akcay2007]. However, despite such well-established hypotheses, no modelling investigates conditions under which pre-copulatory or post-copulatory inbreeding avoidance, or both strategies, are likely to evolve, or predicts how such evolution will feed back to affect underlying evolution of polyandry.

Meanwhile, studies of diverse species have found evidence for female inbreeding avoidance enacted through pre-copulatory or post-copulatory routes, or both. For example, pre-copulatory inbreeding avoidance has been documented in weevils, birds, and butterflies [e.g., @Kuriwada2011; @Kingma2013; @Fischer2015], and post-copulatory inbreeding avoidance has been documented in crickets [e.g., @Simmons2006; @Bretman2009]. Evidence of both pre-copulatory and post-copulatory inbreeding avoidance has accumulated for some well-studied species, including Trinidadian guppies [*Poecilia reticulata*; @Gasparini2011; @Daniel2015] and house mice [*Mus domesticus*; @Potts1991; @Firman2015]. While most studies consider either pre-copulatory or post-copulatory inbreeding avoidance in isolation, a small number of studies consider both. Accordingly, @Liu2014 found evidence of pre-copulatory, but not post-copulatory, inbreeding avoidance in cabbage beetles (*Colaphellus bowringi*). Other studies found no evidence of pre-copulatory or post-copulatory inbreeding avoidance despite strong inbreeding depression [e.g., @Reid2014; @Reid2015]. These studies demonstrate that diverse combinations of pre-copulatory and post-copulatory inbreeding avoidance occur in nature. However, there is as yet no modelling framework that predicts what combinations should be favoured by selection when both pre-copulatory and post-copulatory inbreeding avoidance, and polyandry, can evolve. Such a framework would allow prediction of the form of mating system evolution across different systems that experience inbreeding depression and allow existing empirical studies to be compared to model predictions.

>>>>

Inbreeding avoidance has been widely hypothesised to evolve through pre-copulatory and post-copulatory mechanisms as a consequence of inbreeding depression in offspring viability and female multiple mating (polyandry). However, no theoretical framework exists from which to derive hypothese concerning joint evolution of pre-copulatory and post-copulatory inbreeding avoidance. Evolution of post-copulatory inbreeding avoidance necessitates polyandry, potentially leading to complex evolutionary dynamics among inbreeding strategies and female mating frequency, each of which might be affected by direct negative costs. Hence, we use individual-based modelling to quantify evolution of interacting pre-copulatory and post-copulatory inbreeding avoidance and associated polyandry across multiple cost combinations given severe inbreeding depression. We find that evolution of post-copulatory inbreeding avoidance increases selection for polyandry, that selection for costly inbreeding strategies was weak or negligible given a non-costly alternative strategy. Established pre-copulatory inbreeding avoidance generally precluded evolution of polyandry and therefore inbreeding avoidance, but established post-copulatory inbreeding aviodance did not preclude pre-copulatory inbreeding avoidance from evolving. We thereby provide a general framework for understanding evolution of pre-copulatory and/or post-copulatory inbreeding avoidance across systems, and conclude that evolution of inbreeding avoidance phenotypes and associated polyandry are predicted to interact and be strongly affected by direct costs.

@@

<!--- Note that there are twice as many butterfly species and about eight times as many weevil species as there are bird species. I think that weevils and butterflies can be on at least equal footing with birds in terms of generality --->

Inbreeding, defined as reproduction between relatives, often greatly reduces the fitness of resulting inbred offspring [termed 'inbreeding depression'; @Charlesworth1999; @Keller2002; @Charlesworth2009]. Such strong inbreeding depression is widely hypothesised to drive evolution of inbreeding avoidance, which can be enacted through multiple reproductive decisions [@Parker1979; @Parker2006; @Pusey1996; @Szulkin2012]. From the perspective of a focal breeding female, inbreeding avoidance might be achieved by avoiding mating with related males (i.e., pre-copulatory inbreeding avoidance), and by biasing fertilisation towards unrelated males after mating has occurred (i.e., post-copulatory inbreeding avoidance). Evolution of post-copulatory inbreeding avoidance requires that females are polyandrous, defined as mating with multiple males during a single reproductive bout. Further, such polyandry could itself evolve because it allows females to mate with additional unrelated males following an initial mating with a relative, thereby facilitating pre-copulatory inbreeding avoidance even without any post-copulatory female choice or otherwise biased fertilisation (i.e., given a 'fair raffle' among sperm provided by females' mates). Polyandry can thereby simultaneously allow females to mate with less closely related males and create the opportunity for further inbreeding avoidance enacted through active post-copulatory choice. Polyandry, which might not increase female reproductive success and therefore requires explanation [@Parker2013], has consequently been widely hypothesised to evolve as an adaptation to allow females to avoid inbreeding and thereby increase offspring fitness [@Zeh1997; @Jennions2000; @Tregenza2002]. However, there is surprisingly little theory or modelling that critically examines such widely-invoked hypotheses, and hence examines the conditions under which pre-copulatory or post-copulatory inbreeding avoidance, or both strategies, are predicted to evolve, or how evolution of such strategies of inbreeding avoidance might feed back to affect underlying evolution of polyandry.

Despite a lack of non-verbal theory and resulting predictions, numerous studies on diverse species have found evidence for female inbreeding avoidance. However, only a few such studies have determined whether inbreeding avoidance is enacted through pre-copulatory or post-copulatory mechanisms. Among these few studies, pre-copulatory inbreeding avoidance has been documented in weevils, birds, and butterflies [e.g., @Kuriwada2011; @Kingma2013; @Fischer2015], and post-copulatory inbreeding avoidance has been documented in crickets [e.g., @Simmons2006; @Bretman2009]. An even smaller number of studies comprise sufficient evidence for mechanisms of both pre-copulatory and post-copulatory inbreeding avoidance in two well-studied species, including Trinidadian guppies [*Poecilia reticulata*; @Gasparini2011; @Daniel2015] and house mice [*Mus domesticus*; @Potts1991; @Firman2015]. A lack of both pre-copulatory and post-copulatory inbreeding has been found in fruit flies [e.g., @Ala-Honkola2011; @Tan2012], and might be reasonably inferred where no evidence of inbreeding avoidance exists [e.g., @Reid2014; @Reid2015]. To our knowledge, only @Liu2014 consider both pre-copulatory and post-copulatory inbreeding avoidance jointly, finding evidence of pre-copulatory, but not post-copulatory, inbreeding avoidance in cabbage beetles (*Colaphellus bowringi*). Taken together, all of these studies demonstrate that diverse combinations of pre-copulatory and post-copulatory inbreeding avoidance occur in nature. However, there is as yet no framework that predicts what combinations should be favoured by selection when polyandry and pre-copulatory and post-copulatory inbreeding avoidance can all potentially evolve. Consequently, there is no framework that allows the diversity of existing empirical results to be rationalised or interpreted, and no clear hypotheses that could be tested with these studies through comparative analyses or that could be applied to future empirical studies of individual systems.

<!--- LEFT OFF HERE ------------------------------------------------------->
